Countermeasures against co-location attempts in the cloud

### 2015-04-13 paper.tex
merge the paper

\documentclass[conference]{IEEEtran}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{url}
\usepackage{subfigure}
\usepackage{booktabs,threeparttable,multirow}
\usepackage{siunitx}
\usepackage{pgfplots}

@@

\section{Background}\label{sec:background}
The idea of performing cryptanalytic attacks against co-located virtual machines was discussed by Ristenpart, Shacham,
  and Savage in 2009 \cite{ristenpart09}.
They measure co-location success rates in Amazon Web Services (AWS), and suggest some potential attacks against
  co-located VMs.
Zhang \cite{zhang2014cross} et al. presented a cross-VM attack using Ristenpart's attack which was able to recover an El Gamal key across
Xen VMs.
Irazoqui, Inci, Eisenbarth, and Sunar \cite{irazoquis2014wait} then presented a practical, fast, and realistic full key
  recovery against AES across virtual machines, based on Bernstein's \cite{bernstein2005cache} attack.

@@

\subsection{Covert Channel}
Our goal is to be able to discern co-location with an arbitrary instance, but we can begin by attempting to discern
  co-location with another instance under our control.
Ristenpart et al. evaluated co-location by attempting to transmit data across a covert channel between two virtual
  machines --- in their case, by causing seeks and measuring seek times. Our approach is essentially the same.
A ``sender'' process continually reads values from random locations on the disk to send a one, and does nothing to
  send a zero.
A ``receiver'' process would constantly be reading values, and measuring the seek times.
If the processes shared a hard drive (and thus, a host), the receiver process would be expected to observe consistently
  higher seek times while the sender is transmitting ones than while the sender is transmitting zeroes.
Thus, successfully transmitting a message across this secret channel would imply that the virtual machines are
  co-located.

>>>>

\documentclass[conference]{IEEEtran}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{url}
\usepackage{subfigure}
\usepackage{booktabs,threeparttable,multirow}
\usepackage{siunitx}
\usepackage{pgfplots}
\usepackage{graphicx}
\graphicspath{ {images/} }

@@

\section{Background}\label{sec:background}
The idea of performing cryptanalytic attacks against co-located virtual machines was discussed by Ristenpart, Shacham,
  and Savage in 2009 \cite{ristenpart2009cloud}.
They measure co-location success rates in Amazon Web Services (AWS), and suggest some potential attacks against
  co-located VMs.
Zhang \cite{zhang2014cross} et al. presented a cross-VM attack using Ristenpart's attack which was able to recover an El Gamal key across
Xen VMs.
Irazoqui, Inci, Eisenbarth, and Sunar \cite{irazoqui2014wait} then presented a practical, fast, and realistic full key
  recovery against AES across virtual machines, based on Bernstein's \cite{bernstein2005cache} attack.

@@

\subsection{Network Cartography}
In this section, we create a “map” of instance placements by recording the private IP address of the instances after placing them.
Our hope is this will give insight into how target instances are placed.
This will help in the attempt to establish co-residence with these target instances.

From published information about EC2, we know that each availability zone has its own dedicated hardware. 
Using this and the cartography information in Ristenpart, we hypothesize that each zone will have its own 
range of internal IP addresses. 
We also hypothesize that instances of the same type would fall in the same internal IP address ranges. 
If these hypotheses are correct, an attacker could achieve co-location much more easily by finding the internal 
IP address of the target and launching instances with similar parameters. 
Amazon’s DNS service can be used to obtain this address from the public IP address of the target.

We test our theory by launching a number of instances with different types in each availability zone. 
We then attempt to create a set of rules that describe the likely availability zone and instance type of a given 
private IP address.  
With the data collected, we create a “map” of IP addresses and instance parameters that can be used to determine 
a target’s parameters.

All instances were launched in US West 2 zones a, b, and c. The t2.micro, small, and medium instance types 
were used. 
Of the 1011 instances launched, all fell into the same /16 prefix (172.31.0.0). 
Instances were launched 
in groups of 3 or 5 of each type. 
We observed that instances of the same type launched sequentially had consecutive IP addresses, which is called 
sequential placement locality.

While it was fairly likely that instances launched in the same zone would be placed together based on previous 
work and knowledge of the inner workings of EC2, we were surprised to see that different instances types were launched 
very close to each other (within the same /24 prefix). 
Of the 48 /24 prefixes observed across the three availability zones, the vast majority contained more than one 
instance type. 
Additionally, a small number of m3.medium, large, and xlarge instances were launched which all fell into the same 
/20 prefix as other instances in that zone.

The following plots show the results of the data collected. 
IP addresses are represented by their integer representation minus 2887712768, which represents the first address 
(172.31.0.0). 
Zone 2c occupied the 172.31.0.0/20 prefix, Zone 2a occupied 172.31.16.0/20, and Zone 2b occupied 172.31.32.0/20. 
In (this figure), the zones are numbered one to three, with a being one.
%\begin{figure}[h]
%\includegraphics[scale=.5]{zoneIP}
%\caption{Plot of private IP address vs availability zone}
%\end{figure}
In (these figures), the distribution of instance types in each zone is showed. The instances are numbered from 
one to three, for t2.micro, t2.small, and t2.medium instances respectively.
%\begin{figure}[h]
%\includegraphics[scale=.5]{2cTypeIP}
%\caption{Plot of private IP address vs instance type in Zone 2c}
%\end{figure}
%\begin{figure}[h]
%\includegraphics[scale=.5]{2aTypeIP}
%\caption{Plot of private IP address vs instance type in Zone 2a}
%\end{figure}
%\begin{figure}[h]
%\includegraphics[scale=.5]{2bTypeIP}
%\caption{Plot of private IP address vs instance type in Zone 2b}
%\end{figure}
Our goal is to be able to discern co-location with an arbitrary instance, but we can begin by attempting to discern
  co-location with another instance under our control.
Ristenpart et al. evaluated co-location by attempting to transmit data across a covert channel between two virtual
  machines --- in their case, by causing seeks and measuring seek times. Our approach is essentially the same.
A ``sender'' process continually reads values from random locations on the disk to send a one, and does nothing to
  send a zero.
A ``receiver'' process is constantly reading values, and measuring the seek times.
If the processes shared a hard drive (and thus, a host), the receiver process would be expected to observe consistently
  higher seek times while the sender is transmitting ones than while the sender is transmitting zeroes.
Thus, successfully transmitting a message across this secret channel would imply that the virtual machines are
  co-located.
